---
title: "wk1"
output: html_document
date: "2024-01-25"
---

#We start with Loading the packages for reading the raster data

```{r}
library(sf)
library(terra)
library(raster)
```
#Reading the shape file of Landsat polygons of different Land Classification

```{r}
bare_earth <- st_read("Data/shapefile-landsat/bare_earth_Polygon.shp")
```

```{r}
grass <- st_read("Data/shapefile-landsat/Grass_Polygon.shp")
forest <- st_read("Data/shapefile-landsat/Forest_Polygon.shp")
urban <- st_read("Data/shapefile-landsat/urban_Polygon.shp")
high_urban <- st_read("Data/shapefile-landsat/high_urban_Polygon.shp")
```
**we do not need to add the shapefiles for sentinel as they are in the same area not different**



```{r}
#Landsat equivalent
bands <- c("1","2", "3", "4", "5", "6", "7")

sentinel <- rast("Data/sentinel-tiff/LC09_L2SP_175083_20231206_20231207_02_T1.tif")

#names(sentinel) <- bands
  
landsat<-rast("Data/landsat-tiff/subset_LC09_L2SP_175083_20231206_20231207_02_T1.tif")

names(landsat) <- bands
```

Here i have chosen to use terra, this is because i want to extract the pixel values from within the polygons and at the moment stars will only permit aggregation - e.g. the mean of the pixels in the polygons. 

To use a vector layer in terra it needs to be in a SpatVector (in terra the raster is a SpatRaster)…

```{r}
urban <- vect(urban)
```

```{r}
bare_earth <- st_read("Data/shapefile-landsat/bare_earth_Polygon.shp") %>%
  vect()
grass <- st_read("Data/shapefile-landsat/Grass_Polygon.shp")%>%
    vect()
forest <- st_read("Data/shapefile-landsat/Forest_Polygon.shp")%>%
    vect()
urban <- st_read("Data/shapefile-landsat/urban_Polygon.shp")%>%
    vect()
high_urban <- st_read("Data/shapefile-landsat/high_urban_Polygon.shp")%>%
    vect()
```
Check the CRS

```{r}
crs(landsat)
crs(sentinel)

# reproject landsat
landsat <- project(landsat, sentinel)
```
```{r}
library(tidyverse)
```
Now let’s pull out those values, and get the mean and standard deviation, starting with urban from sentinel data…

```{r}
# extract the values from sentinel for urban landcover
sen_urban<- terra::extract(sentinel, urban, progress = F)%>%
  # make to tibble
  as_tibble()%>%
  # columns are bands so pivot all the bands (2 to 7) longer
  # giving a column of band number and another of band values
  pivot_longer(cols = 2:7, 
               names_to="bands", 
               values_to="band_values")%>%
  # add a column to identify this as sentinel
  add_column(sensor="sentinel")%>%
  # add column to identify as urban
  add_column(land="urban")
```

```{r}
band_fun <- function(sensor, landcover) {
  col_sensor <- deparse(substitute(sensor))
  col_land <- deparse(substitute(landcover))

  sen_urban<- terra::extract(sensor, landcover, progress = F)%>%
    as_tibble()%>%
    pivot_longer(cols = 2:7, 
               names_to="bands", 
               values_to="band_values")%>%
    add_column(sensor=col_sensor)%>%
    add_column(land=col_land)
                 
}
```

```{r}
band_fun_all_values <- function(sensor, landcover) {
  col_sensor <- deparse(substitute(sensor))
  col_land <- deparse(substitute(landcover))

  sen_urban<- terra::extract(sensor, landcover, progress = F)%>%
    as_tibble()%>%
    pivot_longer(cols = 2:7, 
               names_to="bands", 
               values_to="band_values")
                 
}
```

```{r}
sen_bare <- band_fun(sentinel, bare_earth)
sen_grass<- band_fun(sentinel, grass) 
sen_forest<- band_fun(sentinel, forest) 
sen_high_urban <- band_fun(sentinel, high_urban) 

lsat_urban<- band_fun(landsat, urban)
lsat_bare<- band_fun(landsat, bare_earth)
lsat_grass<- band_fun(landsat, grass)
lsat_forest<- band_fun(landsat, forest)
lsat_high_urban <- band_fun(sentinel, high_urban) 
```

```{r}
sen_lsat <- bind_rows(sen_urban, sen_bare, sen_grass,
                      sen_forest, sen_high_urban,
                      lsat_urban, lsat_bare, lsat_grass,
                      lsat_forest, lsat_high_urban)
```

get the mean (and standard deviation) values for each band per sensor and land cover type:

```{r}
means<- sen_lsat%>%
  group_by(bands, sensor, land)%>%
  summarise(Mean=mean(band_values), Std=sd(band_values))
```
```{r}
p1 <- means %>%
  filter(sensor=="sentinel") %>%
  ggplot(., aes(x = bands, y = Mean,
                col=land))+
  geom_point()+
  geom_line(aes(group = land)) +
  geom_errorbar(aes(ymin = (Mean-Std), ymax = (Mean+Std), width = 0.2))
p1
```

```{r}
p2 <- sen_lsat %>%
  filter(sensor=="sentinel") %>%
ggplot(., aes(x=band_values, group=land, fill=land)) + 
  geom_density(alpha = 0.6)+
#Add a mean vertical line
  geom_vline(data = . %>% group_by(land) %>% summarise(group_mean = mean(band_values)),
             aes(xintercept=group_mean, color = land), linetype="dashed", size=1)
```
```{r}
p2
```

```{r}
install.packages("cowplot")
library(cowplot)
```
```{r}
output<-plot_grid(p1, p2, labels = c('A', 'B'), label_size = 12, ncol=1)

output
```

```{r}
ggsave("spectral_reflectance.pdf", width = 20, height = 20, units = "cm")
```

```{r}
t1<- sen_lsat %>%
           filter(sensor=="sentinel" & land =="urban")%>%
           select(band_values)%>%
           pull()

t2<- sen_lsat %>%
           filter(sensor=="landsat" & land =="urban")%>%
           select(band_values)%>%
           pull()

t.test(t1, t2)
```















