---
title: "wk3-Corrections"
output:
  pdf_document: default
  html_document: default
date: "2024-01-30"
---
##ATMOSPHERIC CORRECTION
CHECKING THE DIGITAL NUMBER

we have taken two imagery for the study area overlapping with each other.

#merging Imagery (LANDSAT 8 & LANDSAT 9)

**LANDSAT 8**

```{r}
library(tidyverse)
library(terra)
library(fs)

# List your raster files excluding band 8 using the patter argument
listlandsat_8<-dir_info(here::here("Data", "week3", "LC08_L2SP_175084_20240115_20240124_02_T1"))%>%
  dplyr::filter(str_detect(path, "[B123456790].TIF")) %>%
  dplyr::select(path)%>%
  pull()%>%
  as.character()%>%
  # Load our raster layers into a stack
  terra::rast()
```

**LANDSAT 9**

```{r}
# List your raster files excluding band 8 using the patter argument
listlandsat_9<-dir_info(here::here("Data", "week3", "LC09_L2SP_175083_20240123_20240124_02_T1"))%>%
  dplyr::filter(str_detect(path, "[1B23456790].TIF")) %>%
  dplyr::select(path)%>%
  pull()%>%
  as.character()%>%
  # Load our raster layers into a stack
  terra::rast()
```

```{r}
m1 <- terra::mosaic(listlandsat_8, listlandsat_9, fun="mean")
```
#Enhancements
Ratioing is the difference between two spectral bands that have a certain spectral response meaning it is easier to identify a certain landscape feature…for example…NDVI

The **Normalised Difference Vegetation Index** is based on the fact that healthy and green vegetation reflects more in the NIR but absorbs in the Red wavelength

```{r}
m1_NDVI <- (m1$LC08_L2SP_175084_20240115_20240124_02_T1_SR_B5 - m1$LC08_L2SP_175084_20240115_20240124_02_T1_SR_B4 ) / (m1$LC08_L2SP_175084_20240115_20240124_02_T1_SR_B5 + m1$LC08_L2SP_175084_20240115_20240124_02_T1_SR_B4)

m1_NDVI %>%
  plot(.)
```
We can reclassify this to pull out certain areas, for example, only where NDVI is equal to or greater than 0.2

```{r}
veg <- m1_NDVI %>%
  # cbind = combine dataframes, or here or values listed
  terra::classify(., cbind(-Inf, 0.2, NA))

veg %>%
  plot(.)
```

#FILTERING
Filtering refers to any kind of moving window operation to our data which can be saved as a separate raster file.

```{r}
# for a 3 by 3 filter on 1 band, w means the window and 3 means a 3 by 3 square.

m1_filter <- terra::focal(m1$LC08_L2SP_175084_20240115_20240124_02_T1_SR_B4, w=3)
```
# TEXTURE
For the purposes of this part of the practical i have made a study area polygon in QGIS that i will use to clip my raster to, in order to reduce processing time.

```{r}
library(sf)

study_area <- st_read("Data/week3/study_area/study_area_wk3.shp")%>%
  st_transform(., 32634)
  
m1_clip<-m1%>%
  terra::crop(., study_area)%>%
  terra::mask(., study_area)
```
```{r}
install.packages("GLCMTextures")
library(GLCMTextures)

scale <-(m1_clip*0.0000275) + -0.2

textures1<- glcm_textures(
  scale$LC08_L2SP_175084_20240115_20240124_02_T1_SR_B4,
  # size of window
  w = c(7,7),
  # levels means divide the data into 4 "bins" e.g. a range of 0-20
  # would be 0-5, 5-10, 10-15,15-20
  n_levels = 4,
  # raster data might not be grater than 0
  # convert it to a discrete number of grey levels (e.g. 4)
  quantization = "equal range",
  #co-occurence (second order) matrix (1,0) = one pixel to the right
  # default is all directions as below
  shift = list(c(1, 0), c(1, 1), c(0, 1), c(-1, 1)),
  # select what we want
  metrics="glcm_homogeneity") 

plot(textures1)
```
#Data fusion

data fusion is appending new raster data to the existing data or making a new raster dataset with different bands…here we can do this with the texture measure we have created (and the original spectral data if you wish). 

```{r}
# we can just use the c function to combine rasters in terra

raster_and_texture <- c(m1_clip, textures1)
```

#PCA

**Principal Component Analysis** is designed to 
reduce the dimensionality of our data.
Transform multi-spectral data into uncorrelated and smaller dataset
Keep most of the original information
The first component will (should) capture most of the variance within the dataset

```{r}
pca <- prcomp(as.data.frame(raster_and_texture, na.rm=TRUE), 
              center=TRUE, 
              scale=TRUE)

summary(pca)
```

If we were just to print pca then this would give us the “loadings” - the covariances/correlations between the original variables and the new principal components. We can also get rotations pca$rotatation which is adjustment of the locations on a plot. A common method is varimax which adjusts the rotations to make loading high or zero. Imagine a 3D plot and trying to move around the plot to line the data up we have maximum variance on the x-axis. 

A benefit of PCA is that it takes advantage of multicollinearity and makes new variables that are not correlated

```{r}
x <- predict(raster_and_texture, pca)
plot(x)
```
#glcm

```{r}
install.packages("glcm")
library(glcm)
library(raster)

band4_raster<-raster::raster(m1$LC08_L2SP_175084_20240115_20240124_02_T1_SR_B4)

glcm <- glcm(band4_raster,
                   window = c(7, 7),
                   #shift=list(c(0,1), c(1,1), c(1,0), c(1,-1)), 
                   statistics = c("homogeneity"))

glcm$glcm_homogeneity %>%
  plot(.)
```








